A Switching Planner for Combined Task and Observation Planning

### 2010-11-23 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
Another macro, and some more sketching for DTP

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\actions(p_i;(T_i))$ to the problem
for each element, $p_i (T_i)$, of each probabilistic term from
$(:init)$. The physics of these actions is as
follows. $\actions(p_i;(T_i))$ can be executed if no
$\actions(p_j;(T_j))$, $j \neq i$, has been executed from the same
probabilistic term, and, either $(\pp{probabilistic}~..p_i~(T_i)..)$
is in the root conjunct, or it occurs in $T_k$ for some executed
$\actions(p_k;(T_k))$.
%%
Executing $\actions(p_i;(T_i))$ in $s$ effects a transition to a
successor state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with
probability $1 - p_i$. Here, $s^{T_i}$ is the union of $s$ and atomic
terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.

@@

In the decision-theoretic mode, the switching planner solves a sensing
problem in an abstract processes defined in terms of the assumptive
actions in the current sequential plan. Therefore, this mode uses a
different utility model, that reflects the value of performing sensing
actions.

We construct a {\em relaxed} $(\pp{:init})$ block by first
constructing a tree which only includes terms from the {\em relaxed}
visitation characterised by the {\em assumptive} actions in the
current sequential plan. For example, if $\pp{R2D2}$'s plan is:

\[
\begin{array}{l}
\actions(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office});\\
(\pp{report-is-in}~\pp{box}~\pp{office})
\end{array}
\]

@@

In the second step, we iteratively add terms from the original
$(\pp{:init})$ block while the resulting state space remains of a
practicable size. Of course, it is expensive to measure size exactly,
thus in practice we estimate it to be $2^{|\propositions^\#|}$, where
$\propositions^\#$ is the set of propositions in the abstraction
induced by the proposed tree. Terms are proposed for addition, because
they are probabilistic, and because they were omitted in the {\em
relaxed} visitation characterised by {\em assumptive} actions in the
sequential plan. Each iteration, we add one candidate term with the
largest number of parent terms.

Terms a proposed sequentially, with preference given to
terms that are closest to, or which otherwise are, the conjunctive
leaves.

The problem posed to the decision-theoretic system is an abstraction
of the problem at hand with a distinct utility model that focuses the
agents activities on assumption verification. The abstraction is
induced according to the executing sequential plan. Specifically, we
first omit all problem variables that are not mentioned in the linear
trace, and then refine that preliminary model, by progressively adding
more problem variables while the resulting state space remains of a
practicable size. Refinement in our approach is a simple and
inexpensive step, guided by the level at which variable symbols occur
in the description of the initial conditions. The abstract process is
also given a utility model that captures the reward (resp. penalty) of
correctly (resp. falsely) in/validating the assumptions made in the
linear trace. In detail, for each assumption on a state variable, we
can have a reward of $1$ for correctly rejecting it, and a penalty of
$-1$ for falsely rejecting it. Decision theoretic planning proceeds in
a more-or-less continual manner until an assumption is rejected, or
until it returns control to the sequential system because it can see
no benefit from performing further actions in the abstraction.

Addressing briefly the specifics of our mobile robot's
architecture. In our scenario the utility model for the POMDP posed to
decision-theoretic planning is sourced from a motivational component
of our robotic architecture. This proposes a non-zero reward
(resp. penalty) for sequential planning with a true (resp. false)
assumption about a state proposition.

>>>>

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
each element, $p_i (T_i)$, of each probabilistic term from
$(:init)$. The physics of these actions is as
follows. $\assumptiveS(i)$ can be executed if no $\assumptiveS(j)$,
$j \neq i$, has been executed from the same probabilistic term, and,
either $(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct,
or it occurs in $T_k$ for some executed $\assumptiveS(k)$.
%%
Executing $\assumptiveS(i)$ in $s$ effects a transition to a
successor state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with
probability $1 - p_i$. Here, $s^{T_i}$ is the union of $s$ and atomic
terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.

@@

In the decision-theoretic mode, the switching planner solves a sensing
problem in an abstract processes defined in terms of the assumptive
actions in the current sequential plan, and the $(\pp{:init})$ block
from the original problem description.

The focus for decision-theoretic planning is on sensing, rather than
the achievement of utility according to the given objectives. The
process in this mode has a different reward model that reflects the
value of performing sensing actions in the context of a trace proposed
for the preceding sequential deliberation. In detail, for each
$\assumptiveS{i}$ action scheduled in the current sequential plan, we
have a {\em dual} $\assumptiveDT{i}$ so that
$\forall\state\in\states$:

\[
\reward(\state, \assumptiveDT{i}) = \bigg\{ \begin{array}{ll}
\$(T_i) & \pp{if}~\;\;T_i \not\subseteq \state \\
-\$(T_i) & \pp{otherwise} \\
\end{array}
\]

\noindent Where $\$(T_i)$ is a positive numeric quantity which
rewards the agent for correctly jettisoning a false assumption, and
otherwise penalises the agent. Propositionally these dual actions have
no preconditions or effects -- I.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  Additionally, for the action,
$\action_i$, whose scheduled execution triggered the switch to the
decision-theoretic mode, we include an action
$\actions.\poss(\action_i)$ with the reward property:

\[
\reward(\state, \actions.\poss(\action_i)) = \bigg\{ \begin{array}{ll}
\$(\poss(\action_i)) & \pp{if}~\;\; \poss(\action_i) \subseteq \state \\
-\$(\poss(\action_i)) & \pp{otherwise} \\
\end{array}
\]

\noindent Execution of either a $\assumptiveDT{i}$
action, or $\actions.\poss(\action_i)$, returns control to the
sequential planner. If a dual is executed, then the sequential planner
must replan with $\bstate_0$ equal to the current belief. Otherwise,
if $\actions.\poss(\action_i)$ is executed, then the current
sequential plan is executed until further sensing is scheduled, or to
completion. In our project the rewards, $\$:2^\propositions\to\Re$,
for the POMDP posed to decision-theoretic planning is sourced from a
motivational component of our robotic architecture.

The starting belief-state for the abstract process is given in terms
of a {\em relaxed} $(\pp{:init})$ block.  This is constructed by first
constructing a tree which only includes terms from the {\em relaxed}
visitation characterised by the {\em assumptive} actions in the
current sequential plan. For example, if $\pp{R2D2}$'s plan is:

\[
\begin{array}{l}
\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office});\\
(\pp{report-is-in}~\pp{box}~\pp{office})
\end{array}
\]

@@

In a second step, we iteratively refine the {\em relaxed} block by
adding terms from the original block while the resulting state space
remains of a practicable size. Of course, it is expensive to measure
size exactly, thus in practice we estimate it to be
$2^{|\propositions^\#|}$, where $\propositions^\#$ is the set of
propositions in the abstraction induced by the proposed tree. Terms
are proposed for addition, because they are probabilistic, and because
they were omitted in the {\em relaxed} visitation characterised by
{\em assumptive} actions in the sequential plan. Each iteration, we
add one candidate term with the largest number of parent terms already
in the candidate tree.
